---
title: "Chapter 2 - Download Data"
author: "meelyn"
date: "2025-02-25"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = FALSE)
```

# 2 Download Data

To download your CTT data, you first will need to create an SQL relational database. Why create a relational database instead of loading the csv files directly? Using a database uses much less memory when combining or cleaning data frames compared to loading data directly into RStudio. 

If you are new to relational databases, you should use [DuckDB](https://duckdb.org/why_duckdb), a simple database management systems (DMBS) that is easy to install and use in RStudio.

```{r}
# activate renv environment
renv::activate()
```

**!NOTE** The DuckDB install will take ~ 30 min. Please allocate time accordingly.
```{r eval=FALSE}
# install duckdb using renv
library(renv)

renv::install(c('httr', 'DBI', 'RPostgres', 'tidyverse', 'rjson', 'dplyr', 'tidyr', 'tibble', 'readr', 'dotenv'))
# renv::install('duckdb', repos = c('https://duckdb.r-universe.dev', 'https://cloud.r-project.org'))
renv::install('duckdb/duckdb-r')
renv::install('cellular-tracking-technologies/celltracktech')
```

## 2.1 Create a DuckDB database

Below is a sample script to create a DuckDB database

```{r eval=FALSE}
library(celltracktech)

# load env file into environment
load_dot_env(file='.env')

# Settings ----------------------------------------------------------------
my_token <- Sys.getenv('API_KEY') # load env variable into my_token
myproject <- "Meadows V2" # this is your project name on your CTT account, here we are using the CTT project 'Meadows V2'

# Create your data directory if it does not exist
outpath <- "./data/meadows/" # where your downloaded files are to go

create_outpath(outpath)

# Connect to Database using DuckDB -----------------------------------------------------
con <- DBI::dbConnect(duckdb::duckdb(), 
                      dbdir = "./data/meadows/meadows2.db", 
                      read_only = FALSE)
```


## 2.2 Download data from the CTT API

You are now connected to your DuckDB database, but so far nothing is in the database. The script below will download your data from the CTT servers and insert the files into your local database.

NOTE! Your Sensor Station must be set to upload data to our servers. If your station does not upload data, skip this block and go to section 2.2.1.
```{r eval=FALSE}
get_my_data(my_token,
            outpath, 
            con, 
            myproject=myproject,
            begin=as.Date("2023-08-01"),
            end=as.Date("2023-12-31"),
            filetypes=c("raw", "node_health")
)
```

You may get this error message:

```{bash}
Error in post(endpoint = endpoint, payload = payload) : 
  Gateway Timeout (HTTP 504).
```
It is fine, just run the `get_my_data()` function again and it should work properly.

### 2.2.1 Create Database from Files on your Computer
If you already have your Sensor Station files on your computer (i.e. your sensor station is not connected to the internet), you can use the code block below to create a database and add those files to it.
```{r}
celltracktech::create_duck(con)
```

## 2.3 Updating the database

Upload the compressed data (i.e. the '.csv.gz' files) into your DuckDB database:
```{r eval=FALSE}
update_db(con, outpath, myproject)
```

## 2.4 Disconnecting from the database

The benefit of using a relational database is that it does not need to be loaded into your computer memory the entire time you are analyzing your data. You can connect to it, filter/clean the data you want, and then disconnect once you are done, freeing up computer memory for more intensive data analysis tasks.

To disconnect from the database, run the following command:
```{r eval=FALSE}
DBI::dbDisconnect(con)
```

## 2.5 Uploading Node Data from the SD Card

**Note! This step is optional! If you are not uploading data from the Node SD cards, you can skip to chapter 3!**

### 2.5.1 Create Nodes directory

To upload Node data from the SD cards, you will need to create a 'nodes' folder in the CTT Project Name folder. For example, we will create a folder in the 'Meadows V2' folder in the './data/meadows/' directory:

```{r eval=FALSE}
create_outpath('./data/meadows/Meadows V2/nodes/')
```

### 2.5.2 Create individual directories for each Node

Then, create a folder for each Node. In the example below, we are creating a folder for Node 3B8845:
```{r eval=FALSE}
create_outpath('./data/meadows/Meadows V2/nodes/3B8845/')
```

### 2.5.3 Upload Node data to your DuckDB database

Remember, we need to re-connect to the DuckDB database to upload the Node data:
```{r eval=FALSE}
con <- DBI::dbConnect(duckdb::duckdb(), 
                      dbdir = "./data/meadows/meadows.duckdb", 
                      read_only = FALSE)

# Import node data into your database
import_node_data(d = con,
                 outpath = outpath,
                 myproject = myproject)

# update the database
update_db(d = con,
          outpath = outpath,
          myproject = myproject,
          fix = FALSE)

# disconnect from the database
DBI::dbDisconnect(con)
```

You have finished uploading data to your database!
