# About the Functions
There is a sub-folder within this repo named "functions" which is full of, well, scripts that contain functions! You'll notice they're often called (via source()) at the top of the example scripts. This loads in the custom functions that I have written to handle CTT data. Ultimately, these will be rolled into an R package.

## Data Manager
### load_data
**Description**  
Loads data  

**Usage**  
```{r eval=FALSE}
load_data(directory_name=NULL, starttime=NULL, endtime=NULL, tags=NULL)  
```
**Arguments**  

* directory_name: the input folder can contain a miscellany of raw downloaded files from the sensor station (beep data, node health, GPS) all in the same folder or subfolders. Zipped folders need to be unzipped, but compressed files do not (i.e. csv.gz files are just fine as they are).  
* starttime: start time in POSIXct  
* endtime: end time in POSIXct  
* tags: a vector of tag IDs  

**Value**  
The function will return a nested list where each item corresponds to:  
1. beep data  
2. node health  
3. GPS  

Within each list item, there is a list for a data frame and the hardware version. Also, a column "v" has been added to each data frame indicating the hardware version.  

## Node Health  
### node_channel_plots  
**Description**  
This function is the "engine" behind the export function. You can run it standalone with the following parameters, but you don't have to if your sole goal is to output image files.  

**Usage**  
```{r eval=FALSE}
node_channel_plots(health, freq, ids, lat=NULL, lon=NULL)  
```
**Arguments**  

* health: the 2nd list item output by the load_data() function  
* freq: the time interval for which you want variables to be summarized  
* ids: a vector of IDs; the ID is of the format "<radio ID>_<Node ID>"  
* lat: latitude (optional to produce day/night shading)
* lon: longitude (optional to produce day/night shading)

**Value**  
The output is a nested list, where the top level is each combination of channel and node, and each item is a list of the following plots:  

1. battery  
2. RSSI  
3. number of check-ins  
4. scaled number of check-ins as line plot over scaled RSSI  
5. box plot of node RSSI  
THE FOLLOWING ONLY FOR V2  
6. latitude  
7. longitude  
8. scaled RSSI  
9. dispersion  

### node_plots
**Description**  
A set of diagnostic plots per node

**Usage**  
```{r eval=FALSE}
node_plots(health, nodes, freq, lat = NULL, lon = NULL)
```
**Arguments**  

* health: the 2nd data frame output by the load_data() function  
* nodes: list of nodes  
* freq: the time interval for which you want variables to be summarized  
* lat: latitude  
* lon: longitude  

**Value**  
The output is a nested list for each node, with the following plots for each:  

1. RSSI  
2. number of check-ins  
3. battery  
THE FOLLOWING ONLY FOR V2  
4. time mismatches  
5. small time mismatches  

### gps_plots  
**Description**  
Plots to visualize some GPS data. ONLY FOR V2 HARDWARE

**Usage**  
```{r eval=FALSE}
gps_plots(gps, freq)  
```
**Arguments**  

* gps: the 3rd data frame from the load_data() function  
* freq: the time interval of summary  

**Value**  
A list of the following plots:  

1. altitude  
2. number of fixes  

### export_node_channel_plots
**Description**  
Export plots of node x channel data  

**Usage**  
```{r eval=FALSE}
export_node_channel_plots(plotlist=NULL,health,freq="1 hour",out_path=getwd(),whichplots = c(3,2,1))
```
**Arguments**  

* plotlist: allows you to pass the output of node_channel_plots() if you prefer  
* health: the 2nd data frame output by the load_data() function  
* freq: the time interval for which you want variables to be summarized  
* out_path: where you want your plots to go  
* whichplots: an index vector of of the available plots  

**Output**  
This outputs a png for each input combination of node and channel.  

### export_node_plots  
**Description**  
Same as above; index for the plots can be chosen from the list under the `node_plots()` description  
**Usage**  
```{r eval=FALSE}
export_node_plots(plotlist = NULL, health,freq,out_path=getwd(), x=2, y=3, z=1)  
```
**Arguments**  
To assign x, y and z, look at the description for `node_channel_plots()` and select those plot index in the order you want them on the page.  

**Output**  
This outputs a png for each input node
