```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = FALSE)
```

# Obtaining Data from Database

## Duckplyr

If you are new to database queries but familiar with `dplyr` and `tidyverse`, we recommend using `duckplyr`. You can find more information about `duckplyr` [here](https://duckdb.org/2024/04/02/duckplyr.html) but the main takeaway is that you can query your database using dplyr phrases and pipes, while also saving memory on loading data.

```{r}
library(celltracktech)

con <- DBI::dbConnect(duckdb::duckdb(), 
                      dbdir = "./data/Meadows V2/meadows.duckdb", 
                      read_only = FALSE)

# load raw data table and find unique tags
unique_tags = tbl(con, 'raw') |> 
  group_by(tag_id) |>
  summarize(num_detect = n()) |>
  select(tag_id, num_detect) |>
  arrange(desc(num_detect)) |>
  collect()

DBI::dbDisconnect(con)

```

## SQL Queries

If you are more comfortable with the SQL syntax, you can use SQL queries to get data from different tables.

This chapter is a quick summary on how to use SQL in R. If you would like more info on how to run different queries, use this tutorial: [https://solutions.posit.co/connections/db/getting-started/database-queries/](https://solutions.posit.co/connections/db/getting-started/database-queries/) 

## List Tables

List the tables in your database. If everything was run correctly, each data file type (raw, blu, node-health, etc.) should be in its own data table.

Remember to reconnect to the database!

```{r}
library(celltracktech)

con <- DBI::dbConnect(duckdb::duckdb(), 
                      dbdir = "./data/Meadows V2/meadows.duckdb", 
                      read_only = FALSE)

# list tables in database
DBI::dbListTables(con)

# disconnect from database
DBI::dbDisconnect(con)

```
## Find unique tags in `raw` table

```{r}
# connect to database
con <- DBI::dbConnect(duckdb::duckdb(), 
                      dbdir = "./data/Meadows V2/meadows.duckdb", 
                      read_only = FALSE)

# list last 10 records in raw
raw = DBI::dbGetQuery(con, "SELECT * FROM raw
                           ORDER BY time DESC
                           LIMIT 10")
raw

# find unique tags in the raw table
unique_tag = dbGetQuery(con,
                        'SELECT tag_id, COUNT(*) AS num_detect
                        FROM raw
                        GROUP BY tag_id
                        ORDER BY num_detect DESC')
DBI::dbDisconnect(con)

```

Note: you do not need to load each table into RStudio. You should only load the tables you need to for your analysis.