---
title: "Chapter 7 - Activity Budget"
author: "meelyn"
date: "2025-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = FALSE)
```

# 7 Activity Budget

When is the animal active and/or inactive? Use the following script to answer research questions like:

* Nesting/Incubation behavior
* Roosting times
* Foraging times
* Mortality detection

## 7.1 Load settings
```{r}
# activate renv
renv::activate()
# load celltracktech library
library(celltracktech)
library(dplyr)
library(ggplot2)

options(digits = 10)

# DEFS
source("R/defs/plot_themes.R")
# TAG
source("R/functions/tag/tag_functions.R")

# Specify the path to your database file
database_file <- "data/meadows/meadows.duckdb"

start_time <- as.POSIXct("2023-09-01 00:00:00",tz = "GMT")
stop_time <- as.POSIXct("2023-09-02 00:00:00",tz = "GMT")
```

## 7.2 Load Tag Node Detection Data
```{r}
# Connect to database
con <- DBI::dbConnect(duckdb::duckdb(), 
                      dbdir = database_file, 
                      read_only = TRUE)

# load raw data table into RStudio
detection_df <- tbl(con, "raw") |> 
  filter(time >= start_time && time <= stop_time) |>
  collect()

# disconnect from the database
DBI::dbDisconnect(con)
```

## 7.3 Tag Activity
```{r}
# selected_tag_id <- "2D4B782D", # SWSP (swamp sparrow) - Power Tag

# select a specific tag, below is a Power Tag on a Northern Waterthrush (NOWA)
selected_tag_id <- '614B661E'

# subset detection_df dataframe to only included the selected_tag_id
tag_dets <- subset.data.frame(detection_df, tag_id == selected_tag_id)

# sort the rows by time, ascending
tag_dets <- tag_dets[order(tag_dets$time, decreasing = FALSE), ]

tag_beep_interval <- 13 # seconds, will need to know of tag type beep intervals

# calculate tag activity - this turns the number of detections into a single value for a 5 min bin for each Node
tag_activity <- calculate_tag_activity(tag_dets, tag_beep_interval)

# calculate average tag activity
avg_tag_act <- calc_avg_activity(tag_activity, start_time, stop_time)

# set start and stop times for the plots (next section)
plot_start_time <- as.POSIXct("2023-09-01 06:00:00", tz = "GMT")
plot_stop_time <- as.POSIXct("2023-09-02 06:00:00", tz = "GMT")
```

### 7.3.1 Scatter Plot of RSSI vs time by Node
```{r}
ggplot(tag_dets) +
  geom_point(aes(x = time, 
                 y = tag_rssi, 
                 colour = node_id), 
             shape=1) +
  xlim(plot_start_time, plot_stop_time) +
  ggtitle(paste("Detections", selected_tag_id)) +
  xlab("Time (UTC)") + 
  ylab("RSSI (dBm)") +
  classic_plot_theme
```
Here we see that this bird was spending most of the time in the vicinity of nodes 3282EC, 339CAE, and 37824B, but then spends a considerable amount of time near node 32941A.

### 7.3.2 Scatter Plot of activity vs time by Node

Next we will quantify the activity level. Activity is an arbitrary variable; it can be movement, roosting, a specific behavior, etc. You will need to define what activity is based on your study.
```{r}
ggplot(tag_activity) +
  geom_point(aes(x = time, 
                 y = abs_act, 
                 colour = node_id), 
             shape=1) +
  ggtitle(paste("Detections", selected_tag_id)) +
  xlab("Time (UTC)") +
  ylab("Activity (Arb. Units)") +
  xlim(plot_start_time, plot_stop_time) +
  classic_plot_theme
```
Here we see the activity level (y-axis) for every 5 min from September 1 to September 2. Each color represents a unique Node ID. From a brief glance, this tag had low to mid activity levels near nodes 3282EC and 37824B, but it is still difficult to discern what is going on in this plot.

### 7.3.3 2D Histogram of Activity over Time

We can summarize the above plot even further by creating a heatmap. The code chunk below creates a heatmap bin plot, which displays the activity for an individual over a period of time, with the color of each bin representing the count of activity (on a log scale), while the bins tell us the activity level.
```{r}

my_breaks <- c(1, 10, 100, 1000, 10000)

ggplot(tag_activity, 
       aes(x = time, 
           y = abs_act)) +
  geom_bin2d(binwidth = c(3600, 1)) +
  xlim(plot_start_time, plot_stop_time) +
  scale_fill_viridis_c(name = "Counts", 
                       trans = "log", 
                       breaks = my_breaks, 
                       labels = my_breaks) +
  xlab("Time (UTC)") +
  ylab("Activity / Hour (Arb. Units)") +  
  classic_plot_theme
```
From this plot, we can see that there are multiple instances of low activity (demonstrated by the yellow bins at activity/hour 1), and fewer instances of high activity (demonstrated by the blue bins at activity/hour 4 and above).

### 7.3.4 2D histogram of activity vs time WITH avg activity

This code chunk makes the same plot as above, but also displays the average activity level as a red line.
```{r}
ggplot(tag_activity, 
       aes(x = time, 
           y = abs_act)) +
 geom_bin2d(binwidth = c(3600, 1)) +
  geom_line(data = avg_tag_act, 
            aes(x = time, 
                y = avg_activity),
            colour = "Red") +
  geom_point(data = avg_tag_act, 
             aes(x = time, 
                 y = avg_activity), 
             colour = "Red") +
  xlim(plot_start_time, plot_stop_time) +
  xlab("Time (UTC)") +
  ylab("Activity / Hour (Arb. Units)") +
  scale_fill_viridis_c(name = "Counts", 
                       trans = "log", 
                       breaks = my_breaks, 
                       labels = my_breaks) +
  classic_plot_theme
```
Here we see the average activity level for this tag is roughly 1 per hour from September 1 to September 2.

### 7.3.5 Avg activity / hour Vs time.

Let's zoom in and isolate the average activity per hour:

```{r}
ggplot(tag_activity) +
  geom_line(data = avg_tag_act, 
            aes(x = time, 
                y = avg_activity),
            colour = "Red") +
  geom_point(data = avg_tag_act, 
             aes(x = time, 
                 y = avg_activity), 
             colour = "Red") +
  xlim(plot_start_time, plot_stop_time) +
  xlab("Time (UTC)") +
  ylab("Activity / Hour (Arb. Units)") +
  scale_fill_viridis_c(name = "Counts", 
                       trans = "log", 
                       breaks = my_breaks, 
                       labels = my_breaks) +
  classic_plot_theme
```
We see that average activity is highest between 6am-12pm but decreases suddenly after that, which is what we would expect for bird activity level in September.

