
```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = FALSE)
```

# 9 Grid Search Analysis

If you are interested in the extent of an animal's movement, use the following grid search analysis.

Like in Chapter 8 - Habitat Use, the grid search analysis divides an area into a grid, and calculates the received signal strength at each node. Then this process is repeated for each time step in a series of detections recorded by the node network.

## 9.1 Loading Settings

### 9.1.1 Activate renv
```{r}
renv::activate()
```

### 9.1.2 Load libraries
```{r}
library(dplyr)
library(duckdb)
library(ggplot2)
```

### 9.1.3 Load functions

```{r}
# DEFS
source("R/defs/plot_themes.R")
# UTILS
source("R/functions/utils/get_time_value.R")
# NDOE
source("R/functions/node/node_functions.R")
# TAG
source("R/functions/tag/load_node_detection_data.R")
# SIDEKICK
source("R/functions/sidekick/load_sidekick_data.R")
# GRID SEARCH
source("R/functions/grid_search/grid_search_functions.R")
```

### 9.1.4 Load settings
```{r}
# Specify the path to your database file
database_file <- "./data/meadows/meadows.duckdb"

# (OPTIONAL) Specify Node time offsets if necessary
node_time_offset_file <- "./data/node_time_offset_8_2_2023.csv"
node_toff_df <- read.csv(node_time_offset_file)

# Specify the tag ID that you want to locate
my_tag_id <- "072A6633"

# Specify the RSSI vs Distance fit coefficients from calibration
a <- -103.0610446987
b <- -60.6023833206
c <- 0.0120558164
rssi_coefs <- c(a, b, c)

# Specify the time range of node data you want to import for this analysis
#   This range should cover a large time window where your nodes were in
#   a constant location.  All node health records in this time window
#   will be used to accurately determine the position of your nodes
node_start_time <- as.POSIXct("2023-08-01 00:00:00", tz = "GMT")
node_stop_time <- as.POSIXct("2023-08-07 00:00:00", tz = "GMT")

# Specify time range of detection data you want to pull from the DB
det_start_time <- as.POSIXct("2023-08-03 00:00:00", tz = "GMT")
det_stop_time <- as.POSIXct("2023-08-04 00:00:00", tz = "GMT")

# Specify a list of node Ids if you only want to include a subset in calibration
# IF you want to use all nodes, ignore this line and SKIP the step below
# where the data frame is trimmed to only nodes in this list
# my_nodes <- c("B25AC19E", "44F8E426", "FAB6E12", "1EE02113", "565AA5B9", "EE799439", "1E762CF3", "A837A3F4", "484ED33B")

# You can specify an alternative map tile URL to use here
my_tile_url <- "https://mt2.google.com/vt/lyrs=y&x={x}&y={y}&z={z}"
```

## 9.2 Load Node Health data from Database
```{r}
# Load from DB
con <- DBI::dbConnect(duckdb::duckdb(), dbdir = database_file, read_only = TRUE)
node_health_df <- tbl(con, "node_health") |>
  filter(time >= node_start_time & time <= node_stop_time) |>
  collect()
DBI::dbDisconnect(con)
```

## 9.3 Get Node Locations
```{r}
# Calculate the average node locations
node_locs <- calculate_node_locations(node_health_df)
# Plot the average node locations
node_loc_plot <- plot_node_locations(node_health_df,
                                     node_locs,
                                     theme = classic_plot_theme)
node_loc_plot
# Write the node locations to a file
export_node_locations("examples/meadows/output/node_locations.csv", node_locs)
# Draw a map with the node locations
node_map <- map_node_locations(node_locs, tile_url = my_tile_url)
node_map
```

## 9.4 Load Station Detection Data
```{r}
# Load from DB
con <- DBI::dbConnect(duckdb::duckdb(), dbdir = database_file, read_only = TRUE)
detection_df <- tbl(con, "raw") |>
  filter(time >= det_start_time & time <= det_stop_time) |>
  filter(tag_id == my_tag_id) |>
  collect()
DBI::dbDisconnect(con)
detection_df <- detection_df %>% mutate(time_value = as.integer(time))
```

## 9.5 Build a Node Grid
```{r}
grid_center_lat <- 38.93664800
grid_center_lon <- -74.9462
grid_size_x <- 500 # meters
grid_size_y <- 800 # meters
grid_bin_size <- 5 # meters
# Create a data frame with the details about the grid
grid_df <- build_grid(
  node_locs = node_locs,
  center_lat = grid_center_lat,
  center_lon = grid_center_lon,
  x_size_meters = grid_size_x,
  y_size_meters = grid_size_y,
  bin_size = grid_bin_size
)
# Draw all of the grid bins on a map
grid_map <- draw_grid(node_locs, grid_df)
grid_map

```

## 9.6 (Optional) Calculate Test Solution

```{r}
test_time <- as.POSIXct("2023-08-03 19:57:50", tz = "GMT")
test_rec_df <- calc_receiver_values(
  current_time = test_time,
  det_window = 60,
  station_tag_df = detection_df,
  node_locs = node_locs,
  node_t_offset = node_toff_df,
  rssi_coefs = rssi_coefs,
  filter_alpha = 0.7,
  filter_time_range = 120
)
print(test_rec_df)

# Find the GridSearch Solution
test_grid_values <- calc_grid_values(grid_df, test_rec_df, rssi_coefs)
solution <- subset(test_grid_values, test_grid_values$value == max(test_grid_values$value))
print(solution)

# Multilateration calculation
reduced_rec_df <- subset.data.frame(test_rec_df, test_rec_df$filtered_rssi >= a)
node_w_max <- reduced_rec_df[reduced_rec_df$filtered_rssi == max(reduced_rec_df$filtered_rssi),]
multilat_fit <- nls(reduced_rec_df$exp_dist ~ haversine(reduced_rec_df$lat,reduced_rec_df$lon,ml_lat,ml_lon),
                      reduced_rec_df,
                      start= list(ml_lat = node_w_max$lat, ml_lon = node_w_max$lon),
                      control=nls.control(warnOnly = T, minFactor=1/65536, maxiter = 100)
                    )
print(multilat_fit)

co <- coef(summary(multilat_fit))

print(paste(co[1,1],co[2,1]))

multilat_result <- c(co[1,1],co[2,1])
test_map <- draw_single_solution(test_rec_df, 
                                 test_grid_values, 
                                 solution, 
                                 multilat_result, 
                                 my_tile_url)
test_map
```

## 9.7 Calculate Track

```{r}
track_df <- calculate_track(
  start_time = "2023-08-03 19:50:45",
  length_seconds = 1050,
  step_size_seconds = 10,
  det_time_window = 60, # Must have detection within this window to be included in position calculation
  filter_alpha = 0.7,
  filter_time_range = 120, # Time range to include detections in filtered value
  grid_df = grid_df,
  detection_df = detection_df,
  node_locs = node_locs,
  node_t_offset = node_toff_df,
  rssi_coefs = rssi_coefs,
  track_frame_output_path = NULL # If NULL no individual frames will be saved
)
print(track_df)
track_map <- map_track(node_locs, track_df, my_tile_url)
track_map
```

## 9.8 (Optional) Compare with Known Track

```{r}
# If you've recorded a test track with the sidekick and want to see how well you
# are able to recreate it you can use the commands below.
sidekick_file_path <- "data/meadows/sidekick/calibration_2023_8_3_t2.csv"
# Get Sidekick data from CSV (note I modified the headers for cleaner names in here)
sidekick_df <- read.csv(sidekick_file_path)
# Correct sidekick time formatting
sidekick_df <- sidekick_df %>% mutate(time_utc = substring(c(sidekick_df$time_utc), 1, 19))
# Add numerical time value column
sidekick_df <- sidekick_df %>% mutate(time_value = get_time_value(sidekick_df$time_utc))
# Trim Sidekick data to the time of the calculated track
sidekick_df <- subset.data.frame(sidekick_df, time_value <= max(track_df$time))

track_error_df <- calc_track_error(sidekick_df, track_df)
print(track_error_df)
print(min(track_error_df$error))
print(max(track_error_df$error))
print(paste("GS Solution Error = ", mean(track_error_df$error), " +/- ", sd(track_error_df$error)))
#print(paste("ML Solution Error = ", mean(track_error_df$ml_error), " +/- ", sd(track_error_df$ml_error)))

compare_map <- map_track_error(node_locs, track_error_df, sidekick_df, my_tile_url)
compare_map
```

## 9.9 Plot Uncertainty Analysis
```{r}
# need legend for different color dots
ggplot() +
  geom_point(data = track_error_df, aes(x = i, y = ml_error), color = "orange") +
  geom_point(data = track_error_df, aes(x = i, y = error)) +
  xlab("Track Point #") +
  ylab("Solution Error (m)") +
  classic_plot_theme

ggplot() +
  #geom_point(data = track_error_df, aes(x = max_rssi, y = ml_error), color = "orange") +
  geom_point(data = track_error_df, aes(x = max_rssi, y = error)) +
  xlab("Max RSSI (dBm)") +
  ylab("Position Error (m)") +
  classic_plot_theme
```